Computing solution space properties of combinatorial optimization problems via generic tensor networks

### 2022-09-02 paper/paper.tex
fix sampling complexity

\begin{corollary}\label{thm:complexexprk}
        Let $G = (V, E)$ be a graph and $\mathcal{I}_k$ be the number of independent sets with size $k$. The independent sets with size $k > \alpha(G)-K$ of can be computed in time $O(K\log K)\cc + O(|V|\sum_{k=\alpha(G)-K+1}^{\alpha(G)}|\mathcal{I}_k|)$.
\end{corollary}
Again, the theorem can be proved by relating it with its set algebra version in \Cref{thm:complexsetk}. Although likely, one cannot collect all configurations represented by a sum-product expression tree into a vector due to its extremely large configuration space,
one can draw samples unbiasedly from it by direct sampling.
\begin{theorem}
    Let $G = (V, E)$ be a graph, its independent sets with size $k > \alpha(G)-K$ of can be unbiasedly sampled in time $\cc + O(|V|^2 M)$, where $M$ is the number of samples.
\end{theorem}
\begin{proof}
Similar to the proof of \Cref{thm:complexexpr}, we first obtain a sum-product tree representation of the configurations in time $\cc$.
Then we show to generate a sample, the cost is $O(|V|^2)$.
The sampling program starts from the root node and descends this tree recursively to the left and right siblings.
If a node has type \texttt{SUM}, the program draws samples from the left and right siblings with a probability decided by the size of each sub-tree and returns the union of samples.
Otherwise, if a node has type \texttt{PROD}, the program draws two sets of samples of equal sizes from its left and right siblings and returns the element-wise multiplication ($\lor^\circ$) of them.
The recursion stops at a \texttt{LEAF} node having size $1$ or a \texttt{ZERO} node having size $0$.
In a sum-product expression tree, the number of configurations of sub-trees can be determined easily as we will show in the following example.
\begin{example}
Let us consider the following sum-product expression tree 
\begin{equation*}
    (\texttt{SUM}, /, (\texttt{PROD}, /, A, B), (\texttt{SUM}, /, C, D)),
\end{equation*}
where sub-trees $A, B, C$ and $D$ can be any of the four types of nodes.
This sum-product expression tree can be represented diagrammatically as the following:

@@

Since the depth of the expression tree is $O(|V|)$ and the cost of each arithmetic operation is also $O(|V|)$, the overall time complexity to generate a sample is $O(|V|^2)$.
Proving the theorem.
\end{proof}

Similarly, if one is only interested in obtaining independent sets with largest $K$ sizes, we have the following corollary.
\begin{corollary}
    Let $G = (V, E)$ be a graph, its independent sets with size $k > \alpha(G)-K$ of can be unbiasedly sampled in time $O(K\log K)\cc + O(|V|^2 M)$, where $M$ is the number of samples.
\end{corollary}

>>>>

\begin{corollary}\label{thm:complexexprk}
        Let $G = (V, E)$ be a graph and $\mathcal{I}_k$ be the number of independent sets with size $k$. The independent sets with size $k > \alpha(G)-K$ of can be computed in time $O(K\log K)\cc + O(|V|\sum_{k=\alpha(G)-K+1}^{\alpha(G)}|\mathcal{I}_k|)$.
\end{corollary}
Again, the theorem can be proved by relating it with its set algebra version in \Cref{thm:complexsetk}. Although likely, one cannot collect all configurations represented by a sum-product expression tree into a vector due to its extremely large configuration space,
one can draw samples unbiasedly from it by direct sampling.
\begin{theorem}
    Let $G = (V, E)$ be a graph, its independent sets with size $k > \alpha(G)-K$ of can be unbiasedly sampled in time $\cc + O(|V||E| M)$, where $M$ is the number of samples.
\end{theorem}
\begin{proof}
Similar to the proof of \Cref{thm:complexexpr}, we first obtain a sum-product tree representation of the configurations in time $\cc$.
Then we show to generate a sample, the cost is $O(|V||E|)$.
The sampling program starts from the root node and descends this tree recursively to the left and right siblings.
If a node has type \texttt{SUM}, the program draws samples from the left and right siblings with a probability decided by the size of each sub-tree and returns the union of samples.
Otherwise, if a node has type \texttt{PROD}, the program draws two sets of samples of equal sizes from its left and right siblings and returns the element-wise multiplication ($\lor^\circ$) of them.
The recursion stops at a \texttt{LEAF} node having size $1$ or a \texttt{ZERO} node having size $0$.
In a sum-product expression tree, the number of configurations of sub-trees can be determined easily as we will show in the following example.
\begin{example}
Let us consider the following sum-product expression tree 
\begin{equation*}
    (\texttt{SUM}, /, (\texttt{PROD}, /, A, B), (\texttt{SUM}, /, C, D)),
\end{equation*}
where sub-trees $A, B, C$ and $D$ can be any of the four types of nodes.
This sum-product expression tree can be represented diagrammatically as the following:

@@

Since the depth of the expression tree is $O(|E|)$ and the cost of each arithmetic operation is $O(|V|)$, the overall time complexity to generate a sample is $O(|V||E|)$.
Proving the theorem.
\end{proof}

Similarly, if one is only interested in obtaining independent sets with largest $K$ sizes, we have the following corollary.
\begin{corollary}
    Let $G = (V, E)$ be a graph, its independent sets with size $k > \alpha(G)-K$ of can be unbiasedly sampled in time $O(K\log K)\cc + O(|V||E| M)$, where $M$ is the number of samples.
\end{corollary}
